A Triple Origin for Twin Blue Stragglers in Close Binaries

### 2019-01-14 BSS7782.tex
new figures

In spite of these specific predictions for the expected properties of
BSs formed from each of the above production mechanisms, many BSs
exist with observed properties that defy these simple scenarios.  For
example, in the old open cluster M67, there lurks a candidate triple
system that is posited to host two BSs
\citep{2001A&A...375..375V,2003AJ....125..810S}.  The observations
suggest that the outer tertiary is itself a BS, with a mass $\sim$ 1.7
M$_{\odot}$ and orbiting the inner binary with a period of $\sim
1188.5$ days \citep{2003AJ....125..810S}.  The inner binary has a
period of only $\sim 1.068$ days \citep{2001A&A...375..375V}, and
hosts a BS of mass $\sim 2.52$ M$_{\odot}$.  In order to form this
system we require at least five stars in order to conserve the total
mass in the system \citep{2011MNRAS.410.2370L}.  As explained in
\citet{2011MNRAS.410.2370L}, this is strongly indicative of a
dynamical origin for the system, and a single direct interaction
involving a binary and a triple that resulted in two separate
collision events is the most probable explanation for its origin
(i.e., a single interaction involving two multiples with two or more
stars is a more likely scenario to produce this system than two
back-to-back direct binary-binary interactions)
\citep{2004MNRAS.350..615G}.  \simon{Maybe add something about the
  curious binary AE Aurigae? - NL2: For sure.  I still don't know
  anything about it... But if I get time over the next few days I will
  do my research and write something.}

@@

\begin{figure}[ht!]
\plotone{fig1.eps}
\caption{Cartoon depiction of our proposed scenario for the formation
  of Binary 7782, specifically mass transfer from an evolved outer
  tertiary companion on to a compact inner binary via a circumbinary
  disk.  The outer tertiary component has mass m$_{\rm 3}$, whereas
  the inner binary components have masses m$_{\rm 1}$ and m$_{\rm 2}$.
  The inner and outer orbital separations are denoted by,
  respectively, a$_{\rm in}$ and a$_{\rm out}$.  The circularization
  radius of the accretion stream is denoted a$_{\rm c}$, as calculated
  via Equation~\ref{eqn:ac}, and marks the mean separation of the
  circumbinary disk.\
  \simon{Should we also indicate the accretion stream onto the two stars? - NL2:  I have now updated this plot to show two circumstellar disks, and the flow on to thiose.  I can edit further, if you recommend it, but it's been tough with the down-sizing.  Maybe if I make things bigger still...  Anyways, happy to try more.}
\label{fig:fig1}}
\end{figure}

@@

\begin{figure}[ht!]
\plotone{fig2.eps}
\caption{Parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane
  allowed for the hypothetical outer tertiary orbit of Binary 7782.
  The solid diagonal black line shows the period corresponding to the
  circularization radius a$_{\rm c}$ for the mass transfer stream
  coming from the outer star (i.e., at the onset of mass transfer), assuming a mass ratio of q $=$ 0.7 and component masses 
  of m$_{\rm 1} =$ 0.9 M$_{\odot}$, m$_{\rm 2} =$ 1.1 M$_{\odot}$ and 
  m$_{\rm 3} =$ 1.4 M$_{\odot}$. \simon{what is the purpose of extending q to 0.1 and
    10? - NL2: I was originally thinking that q changes during MT, but I have now removed the other lines and just left the one corresponding to our test case.} We assume initial component masses of m$_{\rm 1}$ = 1.1
  M$_{\rm \odot}$ and m$_{\rm 2}$ = 0.9 M$_{\rm \odot}$ for the inner
  binary components, and m$_{\rm 3}$ is computed for the outer
  tertiary according to our assumed mass ratio (with our fiducial case
  corresponding to q $=$ 0.7).  We assume completely conservative mass
  transfer for this exercise, and a final mass for the outer tertiary
  of 0.6 M$_{\rm \odot}$ once it has become a WD.  The dashed diagonal
  black line shows a rough criterion for dynamical stability in the
  triple, approximately following \citet{mardling99} (i.e., a$_{\rm
    in} <$ 0.1a$_{\rm out}$ is required for long-term dynamical
  stability in equal-mass co-planar triples).  The vertical solid red
  lines show the outer orbital periods corresponding to the hard-soft
  boundary assuming central velocity dispersions of $\sigma =$ 1, 5
  and 10 km s$^{-1}$.  The vertical dashed black lines show the
  maximum outer orbital period P$_{\rm out}$ for which the outer
  tertiary companion is Roche lobe-filling, assuming a stellar radius
  of R$_{\rm 3} =$ 200 R$_{\odot}$ (which corresponds to the maximum stellar radius reached on the AGB for the range of tertiary masses of interest to us.  The horizontal dashed red line shows the observed
  orbital period for Binary 7782, using its observed orbital period
  and our assumed final inner companion masses (i.e., m$_{\rm 1} =$
  m$_{\rm 2} =$ 1.4 M$_{\odot}$).  Finally, the thick solid horizontal
  red line shows the parameter space for P$_{\rm out}$ allowed after
  considering all of the aforementioned criteria.  \simon{I think that
    the figure needs work. I like putting the parameter space in terms
    of orbital periods, but the q=0.1 and q=10 constraints seems
    not very practical. The red labels are not readable, and I am not
    really sure how to read the figure. What about the alternative in
    the next figure, which shows tertiary mass vs. orbital period? - NL2: I have now edited this figure, removed the other q curves, increased font size, etc., but am okay with it if you want to remove it for something else.}
\label{fig:fig2}}
\end{figure}

%NL2: Can we also plot the radii at the terminal-age RGB and AGB, or some maximum radius during each evolutionary phase?  
%I know this will make the plot busier, but these some more relevant to me for mass transfer than the zero-age RGB and 
%AGB radii.  What do you think?
\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_minimumstablesize.pdf}
\caption{Orbital separation as a function of the mass of the
  Roche-lobe filling outer star. The red curves give the stellar radius r$_{\rm 3}$ at the
  first giant branch (thick curve) and at the AGB (thin curve). The
  blue curves give the orbital separation a$_{\rm RLOF}$ at the moment this giant fills its
  Roche-lobe assuming a circular orbit and a total inner binary mass
  of 2.0\,\MSun. The green curves give, under these conditions, the
  maximum orbital separation for a stable inner binary, or a$_{\rm stable}$.
\label{fig:tertiarymass_vs_size}}
\end{figure}

%NL2: I tried to fill in the text below, but please go through and edit it as you see fit.  Also, we need to decide which figures we are going to keep, since I am guessing there is a limit for ApJL.  This of course relates to the current Figure 2, which you may feel should be removed and would be fine with me.
In Fig.\,\ref{fig:tertiarymass_vs_size} we present several characteristic sizes and orbital separations for the problem at hand.  
The red curves show the stellar radius upon reaching the first giant branch (thick curve) and the asymptotic 
giant branch (thin curves).  The blue curves show the orbital separation at the moment of Roche-lobe overflow for each of the aforementioned giants, 
assuming a circular orbit and a total inner binary mass of $0.9 + 1.1$ M$_{\odot}$ $=$ 2.0 M$_{\odot}$.  The green curves show 
for these assumptions the maximum orbital separation for a stable inner binary pair.

@@

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_t0_N80000_M012MSun1109MSun_a02au_e00_inc9deg.pdf}
\caption{Top view of the simulated triple system in which the outer
  star over-fills its Roche-lobe (represented by 80000\,sph particles
  and a core particle of 0.4\,\MSun (blue bullet). The two companion
  stars are represented as red bullets (to the right).
\label{fig:topview_at_t0}}
\end{figure}

@@

\begin{figure}[ht!]
  \includegraphics[width=0.49\columnwidth]{lagrange_points_and_sph.pdf}
~  \includegraphics[width=0.51\columnwidth]{fig_t600_N80000_M012MSun1109MSun_a02au_e00_inc9deg.pdf}
\caption{Same as Fig.\,\ref{fig:topview_at_t0} but at an age of $t
  \simeq 996days$. Left panel shows the equipotential surface of the
  triple overplotted with the gas distribution, the right panel shows
  just the gas and the stars as bullets.
\label{fig:topview_at_t1000day}}
\end{figure}

>>>>

In spite of these specific predictions for the expected properties of
BSs formed from each of the above production mechanisms, many BSs
exist with observed properties that defy these simple scenarios.  For
example, in the old open cluster M67, there lurks a candidate triple
system that is posited to host two BSs
\citep{2001A&A...375..375V,2003AJ....125..810S}.  The observations
suggest that the outer tertiary is itself a BS, with a mass $\sim$ 1.7
M$_{\odot}$ and orbiting the inner binary with a period of $\sim
1188.5$ days \citep{2003AJ....125..810S}.  The inner binary has a
period of only $\sim 1.068$ days \citep{2001A&A...375..375V}, and
hosts a BS of mass $\sim 2.52$ M$_{\odot}$.  In order to form this
system we require at least five stars in order to conserve the total
mass in the system \citep{2011MNRAS.410.2370L}.  As explained in
\citet{2011MNRAS.410.2370L}, this is strongly indicative of a
dynamical origin for the system, and a single direct interaction
involving a binary and a triple that resulted in two separate
collision events is the most probable explanation for its origin
(i.e., a single interaction involving two multiples with two or more
stars is a more likely scenario to produce this system than two
back-to-back direct binary-binary interactions)
\citep{2004MNRAS.350..615G}.

@@

\begin{figure}[ht!]
\plotone{fig1.eps}
\caption{Cartoon depiction of our proposed scenario for the formation
  of Binary 7782, specifically mass transfer from an evolved outer
  tertiary companion on to a compact inner binary via a circumbinary
  disk.  The outer tertiary component has mass m$_{\rm 3}$, whereas
  the inner binary components have masses m$_{\rm 1}$ and m$_{\rm 2}$.
  The inner and outer orbital separations are denoted by,
  respectively, a$_{\rm in}$ and a$_{\rm out}$.  The circularization
  radius of the accretion stream is denoted a$_{\rm c}$, as calculated
  via Equation~\ref{eqn:ac}, and marks the mean separation of the
  circumbinary disk.
\label{fig:fig1}}
\end{figure}

@@

\begin{figure}[ht!]
\plotone{fig2.eps}
\caption{Parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane
  allowed for the hypothetical outer tertiary orbit of Binary 7782
  before Roche-lobe overflow.  The solid diagonal black line shows the
  period corresponding to the circularization radius a$_{\rm c}$ for
  the mass transfer stream coming from the outer star (i.e., at the
  onset of mass transfer), assuming a mass ratio of q $=$ 0.7 and
  component masses of m$_{\rm 1} =$ 0.9 M$_{\odot}$, m$_{\rm 2} =$ 1.1
  M$_{\odot}$ and m$_{\rm 3} =$ 1.4 M$_{\odot}$.  We assume initial
  component masses of m$_{\rm 1}$ = 1.1 M$_{\rm \odot}$ and m$_{\rm
    2}$ = 0.9 M$_{\rm \odot}$ for the inner binary components, and
  m$_{\rm 3}$ is computed for the outer tertiary according to our
  assumed mass ratio (with our fiducial case corresponding to q $=$
  0.7).  We assume completely conservative mass transfer for this
  exercise, and a final mass for the outer tertiary of 0.6 M$_{\rm
    \odot}$ once it has become a WD.  The dashed diagonal black line
  shows a rough criterion for dynamical stability in the triple,
  approximately following \citet{mardling99} (i.e., a$_{\rm in} <$
  0.1a$_{\rm out}$ is required for long-term dynamical stability in
  equal-mass co-planar triples).  The vertical solid red lines show
  the outer orbital periods corresponding to the hard-soft boundary
  assuming central velocity dispersions of $\sigma =$ 1, 5 and 10 km
  s$^{-1}$.  The vertical dashed black lines show the maximum outer
  orbital period P$_{\rm out}$ for which the outer tertiary companion
  is Roche lobe-filling, assuming a stellar radius of R$_{\rm 3} =$
  200 R$_{\odot}$ (which corresponds to the maximum stellar radius
  reached on the AGB for the range of tertiary masses of interest to
  us.  The horizontal dashed red line shows the observed orbital
  period for Binary 7782, using its observed orbital period and our
  assumed final inner companion masses (i.e., m$_{\rm 1} =$ m$_{\rm 2}
  =$ 1.4 M$_{\odot}$).  Finally, the thick solid horizontal red line
  shows the parameter space for P$_{\rm out}$ allowed after
  considering all of the aforementioned criteria.  
\label{fig:fig2}}
\end{figure}

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_M14MSun_outerorbit.pdf}
  \caption{Giant radius as a function of the mass of the core of the
    Roche-lobe filling outer star (dark blue curve).  Here we adopted
    a donor mass of 1.4\,\MSun, but for an 1.2\,\MSun\, donor curve is
    quite similar (see dotted dark-blue curve).  The red squares in
    the curve show the parameters for which we performed more detailed
    gravitational-hydrodynamical simulations (see \S\,\ref{sims}).
    The horizontal dashed line shows the orbital separation of the
    observed twin blue stragger 7782.  The initial triple in which it
    possibly formed must at least have been dynamically stable. The
    minimal orbital separation for the inner binary for which the
    triple is stable is given by the lower green colored curve.
    Donors which are smaller than about 100\,\RSun\, (dark blue curve
    indicated with a$_{\rm stable}$) result in a dynamically unstable
    triple. The minimal core mass associated with a stable triple is
    then indicated by the left-most vertikcal dotted line.  The
    orbital separation at which the donor star overfills its Roche
    lobe is indicated with the light-blue curve. The top curve (dark
    red) shows an estimate of the final orbit separation of the outer
    star, and therefore of the final orbit of the white dwarf around
    the inner twin blue straggler.  For core masses $\apgt
    0.5$\,\MSun\, the final orbital separataion, after mass transfer,
    is smaller than the initial orbit.  Here we adopted an initial
    inner binary mass of (1.0+0.9)\,\MSun\, and a final twin-blue
    straggler mass of (1.4+1.4)\,\MSun.
\label{fig:tertiarymass_vs_size}}
\end{figure}

In Fig.\,\ref{fig:tertiarymass_vs_size} we present several
characteristic sizes and orbital separations for the problem at hand.
The red curves show the stellar radius upon reaching the first giant
branch (thick curve) and the asymptotic giant branch (thin curves).
The blue curves show the orbital separation at the moment of
Roche-lobe overflow for each of the aforementioned giants, assuming a
circular orbit and a total inner binary mass of $0.9 + 1.1$
M$_{\odot}$ $=$ 2.0 M$_{\odot}$.  The green curves show for these
assumptions the maximum orbital separation for a stable inner binary
pair.

@@

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_BBSS_gas_M14Msun_A010au_t0002.pdf}
\caption{Top view of the simulated triple system in which the outer
  star over-fills its Roche-lobe (represented by 80000\,sph particles
  and a core particle of 0.4\,\MSun (blue bullet). The two companion
  stars are represented as red bullets (to the right).
\label{fig:topview_at_t0}}
\end{figure}

@@

\begin{figure}[ht!]
  \includegraphics[width=0.49\columnwidth]{fig_BBSS_Lgas_M14Msun_A010au_t3360.pdf}
~  \includegraphics[width=0.51\columnwidth]{fig_BBSS_gas_M14Msun_A010au_t3360.pdf}
\caption{Same as Fig.\,\ref{fig:topview_at_t0} but at an age of $t
  \simeq 996days$. Left panel shows the equipotential surface of the
  triple overplotted with the gas distribution, the right panel shows
  just the gas and the stars as bullets.
\label{fig:topview_at_t1000day}}
\end{figure}
